### 2025-02-23 sec/4_experimentsandresults.tex

\section{Experiments and Results}
\label{sec:experiments-and-results}

\subsection{Expert Survey: Indistinguishability of Synthetic Images}
\label{subsec:expert-survey:-indistinguishability-of-synthetic-images}
The expert survey results revealed a striking inability of microscopy experts to consistently distinguish synthetic brightfield microscopy images generated by our diffusion model from real images.
The overall classification accuracy all 11 participants and 30 images was remarkably close to chance level, reaching only 50\%.
This near-random performance strongly indicates the high degree of realism achieved by our synthetic image generation approach.
Figure~\ref{fig:conf_matrix} presents the normalized confusion matrix, visually demonstrating the balanced distribution of correct and incorrect classifications for both real and generated images.
Individual image accuracies varied significantly, ranging from approximately 18\% to 90\%, highlighting the inherent variability within both real and synthetic image sets and the complexity of the discrimination task.
Analysis of expert explanations revealed that experts considered subtle image features related to cell appearance, background texture, edge clarity, and perceived noise or artifacts when attempting to differentiate the images.
Prominent terms like ``cell,'' ``background,'' and ``edge'' indicate a focus on core microscopy image components, while terms such as ``pattern,'' ``perfect,'' ``artifacts,'' ``blurry,'' and ``contrast'' suggested that experts were searching for subtle imperfections or stylistic inconsistencies that might betray a synthetic origin.
However, these cues proved unreliable, ultimately leading to the near-chance level classification accuracy and underscoring the perceptual realism of our diffusion-generated brightfield microscopy images.

\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{images/norm_confusion_matrix}
    \caption{Normalized confusion matrix of expert survey classifications. Overall accuracy is approximately 50\%, demonstrating the significant difficulty for experts in distinguishing real and synthetic brightfield microscopy images.}
    \label{fig:conf_matrix}
\end{figure}

\subsection{Object Detection Performance}
\label{subsec:object-detection-performance}
Table~\ref{tab:model-performances} summarizes the object detection performance, presenting mAP metrics for all trained models across the seven datasets.
Analyzing mAP\@50, we observe that models trained with synthetic data replacing parts of the real data (\textit{scc\_10}, \textit{scc\_30}, \textit{scc\_50}) achieve comparable, and in some cases, slightly improved performance compared to models trained solely on real data (\textit{scc\_real}).
Specifically, the YOLOv8s model trained on the \textit{scc\_30} dataset achieved a mAP\@50 of 0.9043, marginally exceeding the 0.8947 achieved by the same model trained on the \textit{scc\_real} dataset.
Similarly, the RT-DETR-l model trained on the \textit{scc\_10} dataset reached a mAP\@50 of 0.9147, very slightly outperforming the 0.9146 achieved on the \textit{scc\_real} dataset.
These results suggest that synthetic data augmentation can maintain or even slightly enhance detection performance at lower IoU thresholds, potentially due to increased data variability and improved model robustness.

However, when examining performance at higher IoU thresholds, specifically mAP\@75 and mAP\@50:95, a subtle trend of performance decrease merges for models trained with increasing proportions of synthetic data.
This trend is more pronounced for the RT-DETR models, indicating a potentially greater sensitivity of transformer-based architectures to the characteristics of synthetic data.

Qualitative assessment of sample detections (examples in supplementary material) revealed that all models, including those trained with synthetic data, generally performed well in detecting cells under diverse conditions, including overlapping cells and cells at image boundaries.
However, subtle differences in detection precision and confidence scores were observed, warranting further investigation into the fine-grained impact of synthetic data on model behavior.

\begin{table}[h!]
    \centering
    \caption{Detailed comparison of the performance metrics for all evaluated models across the different training datasets.}
    \label{tab:model-performances}
    \resizebox{\linewidth}{!}{%
    \begin{tabular}{l|l|c|c|c}
        \toprule
        Model                         & Dataset             & mAP @50                   & mAP @75                   & mAP @50:95                \\
        \midrule
        \multirow[t]{7}{*}{YOLOv8s}   & \textit{scc\_real} & 0.8947                    & \cellcolor{blue!10}0.8095  & \cellcolor{blue!10}0.6557     \\
                                      & \textit{scc\_10}   & 0.8941                    & 0.8053                     & 0.6470                    \\
                                      & \textit{scc\_30}   & \cellcolor{blue!10}0.9043 & 0.8079                     & 0.6448                    \\
                                      & \textit{scc\_50}   & 0.8932                    & 0.7891                     & 0.6325                    \\
                                      & \textit{scc\_add\_10}   & 0.8949               & 0.8068                     & 0.6502                    \\
                                      & \textit{scc\_add\_30}   & 0.8952               & 0.8115                     & 0.6536                    \\
                                      & \textit{scc\_add\_50}   & 0.8940               & 0.8100                     & 0.6514                    \\
        \midrule
        \multirow[t]{7}{*}{YOLOv8m}   & \textit{scc\_real} & \cellcolor{blue!10}0.9038  & \cellcolor{blue!10}0.8203 & \cellcolor{blue!10}0.6637     \\
                                      & \textit{scc\_10}   & 0.9035                     & 0.8093                    & 0.6558                    \\
                                      & \textit{scc\_30}   & 0.8945                     & 0.8076                    & 0.6462                    \\
                                      & \textit{scc\_50}   & 0.8930                     & 0.8040                    & 0.6456                    \\
                                      & \textit{scc\_add\_10}   & 0.8943                    & 0.8088                     & 0.6550                    \\
                                      & \textit{scc\_add\_30}   & 0.9034                    & 0.8224                     & 0.6606                    \\
                                      & \textit{scc\_add\_50}   & 0.8950                    & 0.8218                     & 0.6588                    \\
        \midrule
        \multirow[t]{7}{*}{YOLOv8x}   & \textit{scc\_real} & \cellcolor{blue!10}0.9038  & \cellcolor{blue!10}0.8120 & \cellcolor{blue!10}0.6639     \\
                                      & \textit{scc\_10}   & 0.9032                     & 0.8055                    & 0.6531                    \\
                                      & \textit{scc\_30}   & 0.9031                     & 0.8085                    & 0.6549                    \\
                                      & \textit{scc\_50}   & 0.8935                     & 0.7961                    & 0.6425                    \\
                                      & \textit{scc\_add\_10}   & 0.8941                    & 0.8090                     & 0.6560                    \\
                                      & \textit{scc\_add\_30}   & 0.8942                    & 0.8212                     & 0.6639                    \\
                                      & \textit{scc\_add\_50}   & 0.8938                    & 0.8195                     & 0.6440                    \\
        \midrule
        \multirow[t]{7}{*}{YOLOv9c}   & \textit{scc\_real} & \cellcolor{blue!10}0.9047  & \cellcolor{blue!10}0.8215 & \cellcolor{blue!10}0.6645     \\
                                      & \textit{scc\_10}   & 0.9042                     & 0.8070                    & 0.6531                    \\
                                      & \textit{scc\_30}   & 0.9042                     & 0.8106                    & 0.6568                    \\
                                      & \textit{scc\_50}   & 0.8935                     & 0.7951                    & 0.6422                    \\
                                      & \textit{scc\_add\_10}   & 0.8954                    & 0.8211                     & 0.6615                    \\
                                      & \textit{scc\_add\_30}   & 0.8952                    & 0.8217                     & 0.6605                    \\
                                      & \textit{scc\_add\_50}   & 0.8947                    & 0.8102                     & 0.6594                    \\
        \midrule
        \multirow[t]{7}{*}{YOLOv9e}   & \textit{scc\_real} & 0.9037                     & \cellcolor{blue!10}0.8201 & 0.6629     \\
                                      & \textit{scc\_10}   & \cellcolor{blue!10}0.9041  & 0.8141                    & \cellcolor{blue!10}0.6650 \\
                                      & \textit{scc\_30}   & 0.8946                     & 0.8069                    & 0.6485                    \\
                                      & \textit{scc\_50}   & 0.8938                     & 0.8037                    & 0.6453                    \\
                                      & \textit{scc\_add\_10}   & 0.8950                    & 0.8188                     & 0.6623                    \\
                                      & \textit{scc\_add\_30}   & 0.8952                    & 0.8209                     & 0.6638                    \\
                                      & \textit{scc\_add\_50}   & 0.9031                    & 0.8188                     & 0.6651                    \\
        \midrule
        \multirow[t]{7}{*}{RT-DETR-l} & \textit{scc\_real} & 0.9146                     & \cellcolor{blue!10}0.8169  & \cellcolor{blue!10}0.6614     \\
                                      & \textit{scc\_10}   & \cellcolor{blue!10}0.9147  & 0.8071                    & 0.6574                    \\
                                      & \textit{scc\_30}   & 0.9017                     & 0.7780                    & 0.6240                    \\
                                      & \textit{scc\_50}   & 0.9036                     & 0.7843                    & 0.6298                    \\
                                      & \textit{scc\_add\_10}   & 0.9146                    & 0.8064                     & 0.6526                    \\
                                      & \textit{scc\_add\_30}   & 0.9043                    & 0.7962                     & 0.6457                    \\
                                      & \textit{scc\_add\_50}   & 0.9036                    & 0.7932                     & 0.6372                    \\
        \midrule
        \multirow[t]{7}{*}{RT-DETR-x} & \textit{scc\_real} & \cellcolor{blue!10}0.9164  & \cellcolor{blue!10}0.8257 & \cellcolor{blue!10}0.6748     \\
                                      & \textit{scc\_10}   & 0.9144                     & 0.8083                    & 0.6565                    \\
                                      & \textit{scc\_30}   & 0.9032                     & 0.7830                    & 0.6328                    \\
                                      & \textit{scc\_50}   & 0.9045                     & 0.8032                    & 0.6437                    \\
                                      & \textit{scc\_add\_10}   & 0.9133                    & 0.7991                     & 0.6459                    \\
                                      & \textit{scc\_add\_30}   & 0.9136                    & 0.8106                     & 0.6532                    \\
                                      & \textit{scc\_add\_50}   & 0.9012                    & 0.7682                     & 0.6203                    \\
        \bottomrule
    \end{tabular}%
    }
\end{table}